Randomized encrypt-then-MAC

### 2024-06-24 faster-fo-transform/etm-transform/main.tex
Reorganized etm-transform section

\section{Introduction}
In the modular FO transform\cite{hofheinz2017modular}, the authors constructed IND-CCA KEM using two steps. The first step is adding ciphertext integrity to a one-way secure encryption scheme (the \monospace{T} transformation). The second step is constructing the IND-CCA KEM out of the OW-PCVA secure encryption scheme (the \monospace{U} transformation). The \monospace{T} transformation uses two techniques for achieving ciphertext integrity, which the authors called ``\emph{de-randomization}'' and ``\emph{re-encryption}''.

Our current proposal is to replace re-encryption with a message authentication code (MAC), but we maintained the deterministic encryption routine. For clarity we will call this ``\emph{de-randomized encrypt-then-MAC}''. The algorithm is specified in figure \ref{fig:derand-etm}.

\begin{figure}[h]
  \begin{minipage}{.49\textwidth}  % Adjust width as needed
    \begin{algorithm}[H]
      \caption{$\encrypt^\notrand_\etm(\pk, m)$}
      \begin{algorithmic}[1]
        \State $(r, k_\mac) \leftarrow G(m)$
        \State $c \leftarrow \encrypt(\pk, m; r)$
        \State $t \leftarrow \sign(k_\mac, c)$
        \State \Return $(c, t)$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
  \hfill  % Add space between algorithms
  \begin{minipage}{.49\textwidth}  % Adjust width as needed
    \begin{algorithm}[H]
      \caption{$\decrypt^\notrand_\etm(\sk, (c, t))$}
      \begin{algorithmic}[1]
        \State $\hat{m} \leftarrow \decrypt(\sk, c)$
        \State $(\hat{r}, \hat{k}_\mac) \leftarrow G(\hat{m})$
        \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
          \State \Return $\bot$
        \EndIf
        \State \Return $\hat{m}$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
  \caption{Derandomized encrypt-then-MAC}\label{fig:derand-etm}
\end{figure}

For the past week I have been thinking about removing the de-randomization, as well. This means that we no longer derive the pseudorandom coin from the message. If the input $\pke$ has a randomized encryption routine, then the transformed encrypt-then-mac scheme will also be randomized. We call this ``\emph{randomized encrypt-then-MAC}''. The algorithm is specified in figure \ref{fig:rand-etm}

\begin{figure}[h]
    \center
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\encrypt^\rand_\etm(\pk, m)$}\label{alg:rand-etm-encrypt}
            \begin{algorithmic}[1]
                \State $k_\mac \leftarrow G(m)$
                \State $c \leftarrow \encrypt(\pk, m)$
                \State $t \leftarrow \sign(k_\mac, c)$
                \State \Return $(c, t)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\decrypt^\rand_\etm(\sk, (c, t))$}\label{alg:rand-etm-decrypt}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
                \State $\hat{k}_\mac \leftarrow G(\hat{m})$
                \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
                    \State \Return $\bot$
                \EndIf
                \State \Return $\hat{m}$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{Randomized encrypt-then-MAC}\label{fig:rand-etm}
\end{figure}

\section{Security of MAC}
Intuitively, if the unauthenticated ciphertext is tempered with, then the tag should change unpredictably. Since the MAC key is derived from the plaintext message, and the plaintext message is hidden behind the OW-CPA security of the input scheme, it is safe to assume that the adversary does not have information on the MAC key. If the adversary does not have the MAC key, then it will not be able to produce the correct tag. Therefore, the adversary cannot produce valid authenticated ciphertexts without knowing the corresponding plaintext in the first place.

However, formalizing this idea into a security proof is surprisingly difficult. The most popular methods of formal security proof is

\bibliographystyle{plain}
\bibliography{./references.bib}

>>>>

\section{Encrypt-then-MAC transformation}
In this section, we introduce two ``encrypt-then-MAC'' ($\etm$ for short) transformations. For a given input public-key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$, the de-randomized $\etm$ transformation returns a a public-key encryption scheme $\pke^\notrand_\etm = (\keygen, \encrypt^\notrand_\etm, \decrypt^\notrand_\etm)$ such that $\encrypt^\notrand_\etm$ is deterministic. On the other hand, the randomized $\etm$ transformation returns a public-key encryption scheme $\pke_\etm = (\keygen, \encrypt_\etm, \decrypt_\etm)$ whose encryption routine preserves the randomization of the input encryption routine.

The encryption routine of both transformations uses the plaintext $m \in \mathcal{M}_\pke$ to derive a MAC key $k_\mac$, then after using the input encryption routine $\encrypt$ to produce an unauthenticated ciphertext $c \in \mathcal{C}_\pke$, the MAC key is used to compute a tag $t$ on the unauthenticated ciphertext. The tuple $(c, t)$ is the authenticated ciphertext that the transformed encryption routine outputs. The decryption routine of both transformations first decrypt the unauthenticated ciphertext to recover the plaintext, then uses the decryption to derive a MAC key, which is used to verify the tag.

Because the tag is computed on the unauthenticated ciphertext, if the unauthenticated ciphertext is tempered with, then the tag must be recomputed. Because of the one-way security of the underlying encryption scheme, it is safe to assume that no adversary can use the unauthenticated ciphertext to recover the message, which means that producing valid tags for new message constitutes a forgery attack. Therefore, mounting any meaningful chosen ciphertext attacks is at least as hard as producing a forgery against the underlying $\mac$ scheme.

The transformed routines $(\encrypt_\etm, \decrypt_\etm)$ and $(\encrypt^\notrand_\etm, \decrypt^\notrand_\etm)$ are described in figure \ref{fig:rand-etm} and \ref{fig:derand-etm} respectively.

\begin{figure}[h]
    \center
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\encrypt_\etm(\pk, m)$}\label{alg:rand-etm-encrypt}
            \begin{algorithmic}[1]
                \State $k_\mac \leftarrow G(m)$  % Need to distinguish from the other G
                \State $c \leftarrow \encrypt(\pk, m)$
                    \Comment{If $\encrypt$ is randomized, then $\encrypt_\etm$ is randomized}
                \State $t \leftarrow \sign(k_\mac, c)$
                \State \Return $(c, t)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\decrypt_\etm(\sk, (c, t))$}\label{alg:rand-etm-decrypt}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
                \State $\hat{k}_\mac \leftarrow G(\hat{m})$
                \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
                    \State \Return $\bot$
                \EndIf
                \State \Return $\hat{m}$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{
        $\etm$ transformation.
    }\label{fig:rand-etm}
\end{figure}

\begin{figure}[h]
  \begin{minipage}{0.49\textwidth}  % Adjust width as needed
    \begin{algorithm}[H]
      \caption{$\encrypt^\notrand_\etm(\pk, m)$}
      \begin{algorithmic}[1]
        \State $k_\mac \leftarrow G(m)$
        \State $r \leftarrow H(m)$
        \State $c \leftarrow \encrypt(\pk, m; r)$
        \State $t \leftarrow \sign(k_\mac, c)$
        \State \Return $(c, t)$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
  \hfill  % Add space between algorithms
  \begin{minipage}{.49\textwidth}  % Adjust width as needed
    \begin{algorithm}[H]
      \caption{$\decrypt^\notrand_\etm(\sk, (c, t))$}
      \begin{algorithmic}[1]
        \State $\hat{m} \leftarrow \decrypt(\sk, c)$
        \State $\hat{k}_\mac \leftarrow G(\hat{m})$
        \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
          \State \Return $\bot$
        \EndIf
        \State \Return $\hat{m}$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
  \caption{Derandomized encrypt-then-MAC}\label{fig:derand-etm}
\end{figure}

% \bibliographystyle{plain}
% \bibliography{./references.bib}
